---
title: "RM2C2 Sample Data Cleaning and QA Report"
format:
  html:
    toc: true
    df-print: paged
params:
  raw_dir: "data_sample"
  output_dir: "output_sample"
execute:
  echo: true
  warning: false
  message: false
---

## Cleaning Workflow

This report follows the same step-by-step pipeline each time it is run.

1. Import raw data from the sample folders.
   Read `.txt` files from `Survey`, `Symbol Search`, `Dot Memory`, and `Color Shapes`.

2. Parse the raw files into structured tables.
   Survey files become one row per visit.
   Cognitive task files become one row per trial.

3. Extract visit metadata from each filename.
   Capture `user_id`, `session`, visit date, file timestamp, and a stable `visit_id`.

4. Standardize data fields.
   Normalize IDs and keep the raw sources aligned to a common visit-level structure.

5. Clean the survey data.
   Keep visit-level metadata such as start and end times, exit status, device information, and self-report answers.

6. Score and summarize `Symbol Search`.
   Use `score_symbol_search()` and `summary_symbol_search()` to create one visit-level row.

7. Score and summarize `Dot Memory`.
   Use `score_dot_memory()` and `summary_dot_memory()` to create one visit-level row.

8. Score and summarize `Color Shapes`.
   Use `score_color_shapes()` and `summary_color_shapes()` to create one visit-level row.

9. Stitch all visit-level data together.
   Join survey and task summaries using `user_id`, `session`, and `visit_id`.

10. Generate the final cleaned dataset.
    Write the cleaned visit-level CSV into `output_sample/`.

11. Run red-flag quality checks.
    Check for missing tasks, unexpected trial counts, survey issues, response-time outliers, low-accuracy patterns, and duplicate visit rows.

12. Export quality-control outputs.
    Write red-flag CSV files for visit-level review and participant-level summary.

13. Render the final report.
    Produce an HTML file that documents the run, outputs, and records needing review.

## Researcher Command

Run this from the project root:

```bash
Rscript -e "library(rmarkdown); render('clean_sample_data.qmd', output_format='html_document', params=list(raw_dir='data_sample', output_dir='output_sample'))"
```

```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(readr)
library(stringr)
library(tibble)
library(knitr)
```

## Run Settings

```{r}
raw_dir <- params$raw_dir
output_dir <- params$output_dir
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

timestamp <- format(Sys.time(), "%Y-%m-%d-T%H-%M-%S")

# The package is not installed in this repo-only workflow, so provide a safe
# fallback for sourced scoring functions that call packageVersion("RM2C2").
packageVersion <- function(pkg) {
  if (identical(pkg, "RM2C2")) {
    return(numeric_version("0.0.0.9000"))
  }
  utils::packageVersion(pkg)
}

source(file.path("R", "add_to.R"))
source(file.path("R", "mult_by.R"))
source(file.path("R", "distance.R"))
source(file.path("R", "SDT_adj.R"))
source(file.path("R", "score_symbol_search.R"))
source(file.path("R", "summary_symbol_search.R"))
source(file.path("R", "score_dot_memory.R"))
source(file.path("R", "summary_dot_memory.R"))
source(file.path("R", "score_color_shapes.R"))
source(file.path("R", "summary_color_shape.R"))
source(file.path("R", "stitch_datasets.R"))
source(file.path("R", "compliance_report.R"))
```

```{r}
expected_trials <- c(
  "Symbol Search" = 18L,
  "Dot Memory" = 3L,
  "Color Shapes" = 10L
)

task_dirs <- c("Survey", "Symbol Search", "Dot Memory", "Color Shapes")
task_files <- map(
  set_names(task_dirs),
  ~ list.files(file.path(raw_dir, .x), pattern = "txt$", full.names = TRUE)
)
```

## Helper Functions

```{r}
clean_name <- function(x) {
  x |>
    str_trim() |>
    str_to_lower() |>
    str_replace_all("[^a-z0-9]+", "_") |>
    str_replace_all("^_|_$", "")
}

extract_filename_meta <- function(path) {
  stem <- basename(path) |>
    str_remove("\\.txt$")
  pieces <- str_split(stem, "_", simplify = TRUE)

  if (str_starts(stem, "game_")) {
    task_name <- str_c(pieces[4:(ncol(pieces) - 4)], collapse = "_") |>
      str_replace_all("_", " ")
    stamp <- pieces[, ncol(pieces) - 3]
  } else {
    task_name <- "Survey"
    stamp <- pieces[, ncol(pieces) - 3] |>
      str_replace_all("x", "")
  }

  visit_date <- str_c(
    pieces[, ncol(pieces) - 2],
    pieces[, ncol(pieces) - 1],
    pieces[, ncol(pieces)],
    sep = "-"
  )

  tibble(
    source_file = path,
    raw_file_name = basename(path),
    task_name = task_name,
    user_id = pieces[, 2],
    session = pieces[, 3],
    visit_date = visit_date,
    file_stamp = stamp,
    visit_id = str_c(pieces[, 2], pieces[, 3], visit_date, sep = "_")
  )
}

parse_kv_string <- function(x, item_sep = ",", kv_sep = ":") {
  parts <- str_split(x, item_sep, simplify = FALSE)[[1]]
  parts <- parts[nzchar(str_trim(parts))]
  keys <- str_extract(parts, paste0("^[^", kv_sep, "]+"))
  values <- str_replace(parts, paste0("^[^", kv_sep, "]+", kv_sep), "")
  tibble(key = clean_name(keys), value = str_trim(values))
}

coerce_columns <- function(df) {
  readr::type_convert(df, na = c("", "NA", "NULL", "NOT SET"))
}

normalize_meta_types <- function(df) {
  for (nm in c(
    "source_file",
    "raw_file_name",
    "task_name",
    "user_id",
    "session",
    "visit_date",
    "file_stamp",
    "visit_id"
  )) {
    if (nm %in% names(df)) {
      df[[nm]] <- as.character(df[[nm]])
    }
  }
  df
}

parse_task_file <- function(path) {
  meta <- extract_filename_meta(path)
  lines <- readLines(path, warn = FALSE)

  rows <- map(
    lines,
    ~ parse_kv_string(.x, item_sep = ",", kv_sep = ":") |>
      tidyr::pivot_wider(names_from = key, values_from = value)
  )

  bind_rows(rows) |>
    mutate(across(everything(), as.character)) |>
    bind_cols(meta[rep(1, length(rows)), c("source_file", "raw_file_name", "visit_date", "file_stamp", "visit_id")]) |>
    coerce_columns() |>
    normalize_meta_types() |>
    mutate(
      user_id = as.character(user_id),
      session = as.character(session),
      task_name = as.character(game_name)
    )
}

parse_survey_file <- function(path) {
  meta <- extract_filename_meta(path)
  lines <- readLines(path, warn = FALSE)

  survey_row <- map(
    lines,
    ~ parse_kv_string(.x, item_sep = "\n", kv_sep = ":")
  ) |>
    bind_rows() |>
    distinct(key, .keep_all = TRUE) |>
    tidyr::pivot_wider(names_from = key, values_from = value)

  if ("id" %in% names(survey_row)) {
    survey_row <- rename(survey_row, survey_user_id = id)
  }
  if ("session" %in% names(survey_row)) {
    survey_row <- rename(survey_row, survey_session = session)
  }

  bind_cols(survey_row, meta) |>
    mutate(
      across(everything(), as.character),
      user_id = coalesce(as.character(survey_user_id), as.character(user_id)),
      session = coalesce(as.character(survey_session), as.character(session))
    ) |>
    select(-any_of(c("survey_user_id", "survey_session"))) |>
    normalize_meta_types()
}

flag_rows <- function(df, flag_type, flag_detail) {
  df |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date,
      flag_type = flag_type,
      flag_detail = flag_detail
    )
}
```

## Parse Raw Files

```{r}
survey_raw <- map_dfr(task_files[["Survey"]], parse_survey_file)
symbol_raw <- map_dfr(task_files[["Symbol Search"]], parse_task_file)
dot_raw <- map_dfr(task_files[["Dot Memory"]], parse_task_file)
color_raw <- map_dfr(task_files[["Color Shapes"]], parse_task_file)
```

```{r}
survey_clean <- survey_raw |>
  transmute(
    user_id = as.character(user_id),
    session = as.character(session),
    visit_id,
    visit_date,
    survey_file = raw_file_name,
    survey_file_stamp = file_stamp,
    pack = pack,
    start_date = start_date,
    start_time = start_time,
    end_date = end_date,
    end_time = end_time,
    exit_status = exit_status,
    exit_screen = exit_screen,
    build_date = build_date,
    application_version = application_version,
    android_version = android_version,
    device_id = device_id,
    device_manufacturer = device_manufacturer,
    device_model = device_model,
    launch_type = launch_type,
    beep_file = beep_file,
    target_beep_time = target_beep_time,
    actual_beep_time = actual_beep_time,
    screen_trail = screen_trail,
    bed_time = bed_time,
    getup_time = getup_time,
    sleep_quality = as.numeric(sleep_quality),
    mindsharp = as.numeric(mindsharp),
    concentrate = as.numeric(concentrate),
    game_distract = as.numeric(game_distract)
  )
```

## Score and Summarize Cognitive Tasks

```{r}
symbol_summary <- symbol_raw |>
  mutate(
    user_id = as.character(user_id),
    session = as.character(session)
  ) |>
  score_symbol_search() |>
  summary_symbol_search(c("user_id", "session", "visit_id"))

dot_summary <- dot_raw |>
  mutate(
    user_id = as.character(user_id),
    session = as.character(session)
  ) |>
  score_dot_memory() |>
  summary_dot_memory(c("user_id", "session", "visit_id"))

color_summary <- color_raw |>
  mutate(
    user_id = as.character(user_id),
    session = as.character(session)
  ) |>
  score_color_shapes() |>
  summary_color_shapes(c("user_id", "session", "visit_id"))
```

## Stitch Cleaned Visit-Level Dataset

```{r}
cleaned_visit_data <- stitch_datasets(
  list(survey_clean, symbol_summary, dot_summary, color_summary),
  common_id = c("user_id", "session", "visit_id")
) |>
  arrange(user_id, session, visit_id)
```

## Red Flag Checks

```{r}
task_counts <- bind_rows(
  symbol_raw |>
    count(user_id, session, visit_id, visit_date, name = "n_trials") |>
    mutate(task_name = "Symbol Search"),
  dot_raw |>
    count(user_id, session, visit_id, visit_date, name = "n_trials") |>
    mutate(task_name = "Dot Memory"),
  color_raw |>
    count(user_id, session, visit_id, visit_date, name = "n_trials") |>
    mutate(task_name = "Color Shapes")
) |>
  mutate(expected_trials = unname(expected_trials[task_name]))

all_visits <- survey_clean |>
  select(user_id, session, visit_id, visit_date) |>
  distinct()

missing_task_flags <- crossing(
  all_visits,
  task_name = names(expected_trials)
) |>
  left_join(task_counts, by = c("user_id", "session", "visit_id", "visit_date", "task_name")) |>
  filter(is.na(n_trials)) |>
  transmute(
    user_id,
    visit_id,
    session,
    visit_date,
    flag_type = "missing_task",
    flag_detail = str_c("No raw file found for ", task_name)
  )

trial_count_flags <- task_counts |>
  filter(n_trials != expected_trials) |>
  transmute(
    user_id,
    visit_id,
    session,
    visit_date,
    flag_type = "unexpected_trial_count",
    flag_detail = str_c(task_name, ": expected ", expected_trials, " trials, found ", n_trials)
  )

survey_flags <- survey_clean |>
  filter(
    is.na(exit_status) |
      exit_status != "NORMAL" |
      is.na(sleep_quality) |
      is.na(mindsharp) |
      is.na(concentrate)
  ) |>
  mutate(
    flag_detail = case_when(
      is.na(exit_status) ~ "Survey exit_status is missing",
      exit_status != "NORMAL" ~ str_c("Survey exit_status is ", exit_status),
      is.na(sleep_quality) ~ "Survey sleep_quality is missing",
      is.na(mindsharp) ~ "Survey mindsharp is missing",
      is.na(concentrate) ~ "Survey concentrate is missing",
      TRUE ~ "Survey contains a missing key field"
    )
  ) |>
  transmute(
    user_id,
    visit_id,
    session,
    visit_date,
    flag_type = "survey_issue",
    flag_detail
  )

rt_flags <- bind_rows(
  symbol_summary |>
    filter(
      SYMBOL_SEARCH.median.RT.all_trials < 250 |
        SYMBOL_SEARCH.median.RT.all_trials > 10000
    ) |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date = str_extract(visit_id, "\\d{4}-\\d{2}-\\d{2}$"),
      flag_type = "rt_outlier",
      flag_detail = str_c(
        "Symbol Search median RT is ",
        round(SYMBOL_SEARCH.median.RT.all_trials, 1),
        " ms"
      )
    ),
  dot_summary |>
    filter(
      DOT_MEMORY.median.RT.all_trials < 250 |
        DOT_MEMORY.median.RT.all_trials > 20000
    ) |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date = str_extract(visit_id, "\\d{4}-\\d{2}-\\d{2}$"),
      flag_type = "rt_outlier",
      flag_detail = str_c(
        "Dot Memory median RT is ",
        round(DOT_MEMORY.median.RT.all_trials, 1),
        " ms"
      )
    ),
  color_summary |>
    filter(
      COLOR_SHAPES.median.RT.all_trials < 250 |
        COLOR_SHAPES.median.RT.all_trials > 10000
    ) |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date = str_extract(visit_id, "\\d{4}-\\d{2}-\\d{2}$"),
      flag_type = "rt_outlier",
      flag_detail = str_c(
        "Color Shapes median RT is ",
        round(COLOR_SHAPES.median.RT.all_trials, 1),
        " ms"
      )
    )
)

accuracy_flags <- bind_rows(
  symbol_summary |>
    filter(SYMBOL_SEARCH.proportion.accurate.trials < 0.50) |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date = str_extract(visit_id, "\\d{4}-\\d{2}-\\d{2}$"),
      flag_type = "low_accuracy",
      flag_detail = str_c(
        "Symbol Search accuracy is ",
        sprintf("%.1f%%", 100 * SYMBOL_SEARCH.proportion.accurate.trials)
      )
    ),
  color_summary |>
    filter(COLOR_SHAPES.CorRec.rate < 0.50) |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date = str_extract(visit_id, "\\d{4}-\\d{2}-\\d{2}$"),
      flag_type = "low_accuracy",
      flag_detail = str_c(
        "Color Shapes correct rejection rate is ",
        sprintf("%.1f%%", 100 * COLOR_SHAPES.CorRec.rate)
      )
    ),
  dot_summary |>
    filter(DOT_MEMORY.prop.perfect.trials == 0) |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date = str_extract(visit_id, "\\d{4}-\\d{2}-\\d{2}$"),
      flag_type = "low_accuracy",
      flag_detail = "Dot Memory has zero perfect trials"
    )
)

duplicate_flags <- bind_rows(
  survey_clean |>
    add_count(user_id, session, visit_id) |>
    filter(n > 1) |>
    distinct(user_id, session, visit_id, visit_date, n) |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date,
      flag_type = "duplicate_visit",
      flag_detail = str_c("Survey has ", n, " rows for this visit")
    )
)

red_flags <- bind_rows(
  missing_task_flags,
  trial_count_flags,
  survey_flags,
  rt_flags,
  accuracy_flags,
  duplicate_flags
) |>
  distinct() |>
  arrange(user_id, visit_date, session, flag_type)

participant_flag_summary <- red_flags |>
  count(user_id, flag_type, name = "n_flags") |>
  arrange(desc(n_flags), user_id, flag_type)

visit_compliance <- compliance_report(cleaned_visit_data, user_id) |>
  arrange(user_id)
```

## Write Output Files

```{r}
clean_path <- file.path(output_dir, str_c("cleaned_visit_data_", timestamp, ".csv"))
flags_path <- file.path(output_dir, str_c("red_flags_", timestamp, ".csv"))
participant_flags_path <- file.path(output_dir, str_c("participant_red_flags_", timestamp, ".csv"))
compliance_path <- file.path(output_dir, str_c("participant_visit_counts_", timestamp, ".csv"))

write_csv(cleaned_visit_data, clean_path)
write_csv(red_flags, flags_path)
write_csv(participant_flag_summary, participant_flags_path)
write_csv(visit_compliance, compliance_path)
```

## Output Files

```{r}
output_manifest <- tibble(
  file_type = c(
    "cleaned_data",
    "visit_red_flags",
    "participant_red_flags",
    "participant_visit_counts"
  ),
  file_path = c(clean_path, flags_path, participant_flags_path, compliance_path)
)

kable(output_manifest)
```

## Sample Overview

```{r}
overview <- tibble(
  dataset = c("Survey", "Symbol Search", "Dot Memory", "Color Shapes", "Final cleaned visits"),
  n_rows = c(
    nrow(survey_clean),
    nrow(symbol_raw),
    nrow(dot_raw),
    nrow(color_raw),
    nrow(cleaned_visit_data)
  ),
  n_unique_participants = c(
    n_distinct(survey_clean$user_id),
    n_distinct(symbol_raw$user_id),
    n_distinct(dot_raw$user_id),
    n_distinct(color_raw$user_id),
    n_distinct(cleaned_visit_data$user_id)
  )
)

kable(overview)
```

## Red Flags To Review

```{r}
if (nrow(red_flags) == 0) {
  cat("No red flags were generated by the default checks.")
} else {
  kable(red_flags)
}
```

## Participant Flag Summary

```{r}
if (nrow(participant_flag_summary) == 0) {
  cat("No participant-level red flag counts were generated.")
} else {
  kable(participant_flag_summary)
}
```
