---
title: "RM2C2 Sample Data Cleaning and QA Report"
format:
  html:
    toc: true
    df-print: paged
params:
  raw_dir: "data_sample"
  output_dir: "output_sample"
execute:
  echo: true
  warning: false
  message: false
---

## Cleaning Workflow

This report follows the same step-by-step pipeline each time it is run.

1. Import raw data from the sample folders.
   Read `.txt` files from `Survey`, `Symbol Search`, `Dot Memory`, and `Color Shapes`.

2. Parse the raw files into structured tables.
   Survey files become one row per visit.
   Cognitive task files become one row per trial.

3. Extract visit metadata from each filename.
   Capture `user_id`, `session`, visit date, file timestamp, and a stable `visit_id`.

4. Resolve interrupted attempts.
   Keep the latest file per visit and hold back visits the app never finished, so a
   restarted session is never scored twice.

5. Standardize data fields.
   Normalize IDs, dates, and clock times, and keep the raw sources aligned to a
   common visit-level structure.

6. Clean the survey data.
   Keep visit-level metadata such as start and end times, exit status, device
   information, self-report answers, and the within-day `moment_surveys` index.

7. Score and summarize `Symbol Search`.
   Use `score_symbol_search()` and `summary_symbol_search()` to create one visit-level row.

8. Score and summarize `Dot Memory`.
   Use `score_dot_memory()` and `summary_dot_memory()` to create one visit-level row.

9. Score and summarize `Color Shapes`.
   Use `score_color_shapes()` and `summary_color_shapes()` to create one visit-level row.

10. Stitch all visit-level data together.
    Join survey and task summaries using `user_id`, `session`, and `visit_id`.

11. Generate the final cleaned dataset.
    Write the cleaned visit-level CSV into `output_sample/`.

12. Run red-flag quality checks.
    Check for missing tasks, unexpected trial counts, survey issues, incomplete and
    superseded attempts, response-time outliers, low-accuracy patterns, and
    duplicate visit rows.

13. Export quality-control outputs.
    Write red-flag CSV files for visit-level review and participant-level summary.

14. Render the final report.
    Produce an HTML file that documents the run, outputs, and records needing review.

## Researcher Command

Run this from the project root:

```bash
Rscript -e "library(rmarkdown); render('clean_sample_data.qmd', output_format='html_document', params=list(raw_dir='data_sample', output_dir='output_sample'))"
```

```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(readr)
library(stringr)
library(tibble)
library(knitr)
```

## Run Settings

```{r}
raw_dir <- params$raw_dir
output_dir <- params$output_dir
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

timestamp <- format(Sys.time(), "%Y-%m-%d-T%H-%M-%S")

# The package is not installed in this repo-only workflow, so provide a safe
# fallback for sourced scoring functions that call packageVersion("RM2C2").
# Read the version from DESCRIPTION so the stamped PACKAGE.VERSION columns match
# the code that actually produced the scores.
rm2c2_version <- numeric_version(
  read.dcf("DESCRIPTION", fields = "Version")[[1]]
)

packageVersion <- function(pkg) {
  if (identical(pkg, "RM2C2")) {
    return(rm2c2_version)
  }
  utils::packageVersion(pkg)
}

source(file.path("R", "add_to.R"))
source(file.path("R", "mult_by.R"))
source(file.path("R", "distance.R"))
source(file.path("R", "SDT_adj.R"))
source(file.path("R", "score_symbol_search.R"))
source(file.path("R", "summary_symbol_search.R"))
source(file.path("R", "score_dot_memory.R"))
source(file.path("R", "summary_dot_memory.R"))
source(file.path("R", "score_color_shapes.R"))
source(file.path("R", "summary_color_shape.R"))
source(file.path("R", "stitch_datasets.R"))
source(file.path("R", "compliance_report.R"))
source(file.path("R", "sample_data_prep.R"))
```

```{r}
# `expected_trials`, `task_dirs`, and `complete_exit_status` come from
# R/sample_data_prep.R, which both this report and the reliability report use.

# Red-flag thresholds, gathered here so they can be retuned without touching
# the check code below.
qc_thresholds <- list(
  rt_min_ms = 250,
  symbol_rt_max_ms = 10000,
  dot_rt_max_ms = 20000,
  color_rt_max_ms = 10000,
  symbol_min_accuracy = 0.50,
  # COLOR_SHAPES.CorRec.rate is corrected recognition (HIT rate - FA rate), not
  # a correct-rejection rate: it runs from -1 to 1 and 0 is chance. Flag only
  # performance at or below chance.
  color_max_corrected_recognition = 0,
  # Dot Memory has 3 trials, so zero *perfect trials* is ordinary (46% of this
  # sample). Zero *perfect dots* across all 9 dots is the rare, real signal.
  dot_max_perfect_dots = 0
)
```

## Helper Functions

Parsing, formatting, and attempt-resolution helpers live in
`R/sample_data_prep.R` so this report and the reliability report clean a visit
the same way. Only the flag helper is local to this report.

```{r}
flag_rows <- function(df, flag_type, flag_detail) {
  df |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date,
      flag_type = flag_type,
      flag_detail = flag_detail
    )
}
```

## Parse Raw Files

`prepare_sample_data()` reads the raw exports, keeps the latest attempt per
visit, and holds back visits the app never finished. A visit can leave more than
one file behind: the app writes what it has when a session is backgrounded or
force-closed, then writes a complete file when the participant returns.

```{r}
prepared <- prepare_sample_data(raw_dir)

task_files <- prepared$task_files
survey_all <- prepared$survey_all
incomplete_surveys <- prepared$incomplete_surveys
survey_raw <- prepared$survey_raw
survey_clean <- prepared$survey_clean
symbol_raw <- prepared$symbol_raw
dot_raw <- prepared$dot_raw
color_raw <- prepared$color_raw
```

## Score and Summarize Cognitive Tasks

```{r}
symbol_summary <- symbol_raw |>
  mutate(
    user_id = as.character(user_id),
    session = as.character(session)
  ) |>
  score_symbol_search() |>
  summary_symbol_search(c("user_id", "session", "visit_id"))

dot_summary <- dot_raw |>
  mutate(
    user_id = as.character(user_id),
    session = as.character(session)
  ) |>
  score_dot_memory() |>
  summary_dot_memory(c("user_id", "session", "visit_id"))

color_summary <- color_raw |>
  mutate(
    user_id = as.character(user_id),
    session = as.character(session)
  ) |>
  score_color_shapes() |>
  summary_color_shapes(c("user_id", "session", "visit_id"))
```

## Stitch Cleaned Visit-Level Dataset

```{r}
cleaned_visit_data <- stitch_datasets(
  list(survey_clean, symbol_summary, dot_summary, color_summary),
  common_id = c("user_id", "session", "visit_id")
) |>
  arrange(user_id, session, visit_id)
```

## Red Flag Checks

```{r}
task_counts <- bind_rows(
  symbol_raw |>
    count(user_id, session, visit_id, visit_date, name = "n_trials") |>
    mutate(task_name = "Symbol Search"),
  dot_raw |>
    count(user_id, session, visit_id, visit_date, name = "n_trials") |>
    mutate(task_name = "Dot Memory"),
  color_raw |>
    count(user_id, session, visit_id, visit_date, name = "n_trials") |>
    mutate(task_name = "Color Shapes")
) |>
  mutate(expected_trials = unname(expected_trials[task_name]))

all_visits <- survey_clean |>
  select(user_id, session, visit_id, visit_date) |>
  distinct()

missing_task_flags <- crossing(
  all_visits,
  task_name = names(expected_trials)
) |>
  left_join(task_counts, by = c("user_id", "session", "visit_id", "visit_date", "task_name")) |>
  filter(is.na(n_trials)) |>
  transmute(
    user_id,
    visit_id,
    session,
    visit_date,
    flag_type = "missing_task",
    flag_detail = str_c("No raw file found for ", task_name)
  )

trial_count_flags <- task_counts |>
  filter(n_trials != expected_trials) |>
  transmute(
    user_id,
    visit_id,
    session,
    visit_date,
    flag_type = "unexpected_trial_count",
    flag_detail = str_c(task_name, ": expected ", expected_trials, " trials, found ", n_trials)
  )

survey_flags <- survey_clean |>
  filter(
    is.na(exit_status) |
      exit_status != "NORMAL" |
      is.na(sleep_quality) |
      is.na(mindsharp) |
      is.na(concentrate)
  ) |>
  mutate(
    flag_detail = case_when(
      is.na(exit_status) ~ "Survey exit_status is missing",
      exit_status != "NORMAL" ~ str_c("Survey exit_status is ", exit_status),
      is.na(sleep_quality) ~ "Survey sleep_quality is missing",
      is.na(mindsharp) ~ "Survey mindsharp is missing",
      is.na(concentrate) ~ "Survey concentrate is missing",
      TRUE ~ "Survey contains a missing key field"
    )
  ) |>
  transmute(
    user_id,
    visit_id,
    session,
    visit_date,
    flag_type = "survey_issue",
    flag_detail
  )

# Visits held out of the cleaned dataset because the app never finished them.
# Reported here so nothing is dropped without a record.
incomplete_flags <- incomplete_surveys |>
  transmute(
    user_id,
    visit_id,
    session,
    visit_date,
    flag_type = "incomplete_session",
    flag_detail = str_c(
      "Held out of cleaned data: exit_status is ",
      exit_status,
      " at ",
      exit_screen
    )
  )

# Superseded partial attempts: a second file exists for the same visit and the
# earlier one was discarded by keep_latest_attempt().
superseded_flags <- survey_all |>
  add_count(visit_id, name = "n_attempts") |>
  filter(n_attempts > 1) |>
  group_by(visit_id) |>
  filter(file_stamp != max(file_stamp)) |>
  ungroup() |>
  transmute(
    user_id,
    visit_id,
    session,
    visit_date,
    flag_type = "superseded_attempt",
    flag_detail = str_c(
      "Earlier attempt discarded (exit_status ",
      exit_status,
      "); kept the later complete file"
    )
  )

rt_flags <- bind_rows(
  symbol_summary |>
    filter(
      SYMBOL_SEARCH.median.RT.all_trials < qc_thresholds$rt_min_ms |
        SYMBOL_SEARCH.median.RT.all_trials > qc_thresholds$symbol_rt_max_ms
    ) |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date = str_extract(visit_id, "\\d{4}-\\d{2}-\\d{2}$"),
      flag_type = "rt_outlier",
      flag_detail = str_c(
        "Symbol Search median RT is ",
        round(SYMBOL_SEARCH.median.RT.all_trials, 1),
        " ms"
      )
    ),
  dot_summary |>
    filter(
      DOT_MEMORY.median.RT.all_trials < qc_thresholds$rt_min_ms |
        DOT_MEMORY.median.RT.all_trials > qc_thresholds$dot_rt_max_ms
    ) |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date = str_extract(visit_id, "\\d{4}-\\d{2}-\\d{2}$"),
      flag_type = "rt_outlier",
      flag_detail = str_c(
        "Dot Memory median RT is ",
        round(DOT_MEMORY.median.RT.all_trials, 1),
        " ms"
      )
    ),
  color_summary |>
    filter(
      COLOR_SHAPES.median.RT.all_trials < qc_thresholds$rt_min_ms |
        COLOR_SHAPES.median.RT.all_trials > qc_thresholds$color_rt_max_ms
    ) |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date = str_extract(visit_id, "\\d{4}-\\d{2}-\\d{2}$"),
      flag_type = "rt_outlier",
      flag_detail = str_c(
        "Color Shapes median RT is ",
        round(COLOR_SHAPES.median.RT.all_trials, 1),
        " ms"
      )
    )
)

accuracy_flags <- bind_rows(
  symbol_summary |>
    filter(SYMBOL_SEARCH.proportion.accurate.trials < qc_thresholds$symbol_min_accuracy) |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date = str_extract(visit_id, "\\d{4}-\\d{2}-\\d{2}$"),
      flag_type = "low_accuracy",
      flag_detail = str_c(
        "Symbol Search accuracy is ",
        sprintf("%.1f%%", 100 * SYMBOL_SEARCH.proportion.accurate.trials)
      )
    ),
  color_summary |>
    filter(COLOR_SHAPES.CorRec.rate <= qc_thresholds$color_max_corrected_recognition) |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date = str_extract(visit_id, "\\d{4}-\\d{2}-\\d{2}$"),
      flag_type = "low_accuracy",
      flag_detail = str_c(
        "Color Shapes corrected recognition (HIT - FA) is ",
        sprintf("%.2f", COLOR_SHAPES.CorRec.rate),
        ", at or below chance"
      )
    ),
  dot_summary |>
    filter(DOT_MEMORY.count.perfect.dots <= qc_thresholds$dot_max_perfect_dots) |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date = str_extract(visit_id, "\\d{4}-\\d{2}-\\d{2}$"),
      flag_type = "low_accuracy",
      flag_detail = "Dot Memory has zero correctly placed dots across all trials"
    )
)

duplicate_flags <- bind_rows(
  survey_raw |>
    add_count(user_id, session, visit_id) |>
    filter(n > 1) |>
    distinct(user_id, session, visit_id, visit_date, n) |>
    transmute(
      user_id,
      visit_id,
      session,
      visit_date,
      flag_type = "duplicate_visit",
      flag_detail = str_c("Survey has ", n, " rows for this visit")
    )
)

red_flags <- bind_rows(
  missing_task_flags,
  trial_count_flags,
  survey_flags,
  incomplete_flags,
  superseded_flags,
  rt_flags,
  accuracy_flags,
  duplicate_flags
) |>
  distinct() |>
  arrange(user_id, visit_date, session, flag_type)

participant_flag_summary <- red_flags |>
  count(user_id, flag_type, name = "n_flags") |>
  arrange(desc(n_flags), user_id, flag_type)

visit_compliance <- compliance_report(cleaned_visit_data, user_id) |>
  arrange(user_id)
```

## Write Output Files

```{r}
clean_path <- file.path(output_dir, str_c("cleaned_visit_data_", timestamp, ".csv"))
flags_path <- file.path(output_dir, str_c("red_flags_", timestamp, ".csv"))
participant_flags_path <- file.path(output_dir, str_c("participant_red_flags_", timestamp, ".csv"))
compliance_path <- file.path(output_dir, str_c("participant_visit_counts_", timestamp, ".csv"))

write_csv(cleaned_visit_data, clean_path)
write_csv(red_flags, flags_path)
write_csv(participant_flag_summary, participant_flags_path)
write_csv(visit_compliance, compliance_path)
```

## Output Files

```{r}
output_manifest <- tibble(
  file_type = c(
    "cleaned_data",
    "visit_red_flags",
    "participant_red_flags",
    "participant_visit_counts"
  ),
  file_path = c(clean_path, flags_path, participant_flags_path, compliance_path)
)

kable(output_manifest)
```

## Sample Overview

```{r}
overview <- tibble(
  dataset = c("Survey", "Symbol Search", "Dot Memory", "Color Shapes", "Final cleaned visits"),
  n_rows = c(
    nrow(survey_clean),
    nrow(symbol_raw),
    nrow(dot_raw),
    nrow(color_raw),
    nrow(cleaned_visit_data)
  ),
  n_unique_participants = c(
    n_distinct(survey_clean$user_id),
    n_distinct(symbol_raw$user_id),
    n_distinct(dot_raw$user_id),
    n_distinct(color_raw$user_id),
    n_distinct(cleaned_visit_data$user_id)
  )
)

kable(overview)
```

## Red Flags To Review

```{r}
if (nrow(red_flags) == 0) {
  cat("No red flags were generated by the default checks.")
} else {
  kable(red_flags)
}
```

## Participant Flag Summary

```{r}
if (nrow(participant_flag_summary) == 0) {
  cat("No participant-level red flag counts were generated.")
} else {
  kable(participant_flag_summary)
}
```
