---
title: "RM2C2 Cognitive Assessment: Reliability for One-Time Use"
format:
  html:
    toc: true
    df-print: paged
params:
  raw_dir: "data_sample"
  output_dir: "output_sample"
  n_boot: 2000
  n_boot_baseline: 4000
  condition_file: ""
execute:
  echo: true
  warning: false
  message: false
---

## Purpose

The M2C2 tasks were designed for ecological momentary assessment: very short
blocks administered many times, where reliability is built up by *averaging
across occasions* rather than by depth at any single one. This study uses them
differently — as **one-time measures** — so the question this report answers is:

> Is a single administration of each task precise enough to rank participants?

That is a different question from the one the EMA reliability literature
answers, and it needs a different coefficient. This report leads with the
single-administration answer and reports the EMA-style coefficients second, with
the reasons they do not apply here.

### Design constraints that shape the analysis

- **Crossover with an intervention.** Occasion-to-occasion differences carry a
  treatment effect, so within-person variance across sessions is not natural
  fluctuation. Session 1 is a clean pre-intervention baseline; sessions 2 and 3
  are not.
- **Sessions are ~15 minutes apart** (median gap 14.9 min, whole protocol inside
  a ~30-minute window). Even without an intervention there is almost no
  day-to-day, sleep, or time-of-day variation for an EMA-style within-person
  coefficient to detect.
- **Baseline is the first occasion per participant, not `session == 1`.**
  Participants 351 and 353 have no session 1 — their sessions run 2, 3, 4. Since
  both completed three sessions like everyone else, that is a session-counter
  offset rather than a missing baseline, so taking the first occasion keeps all
  52 participants instead of dropping to 50.

## Method

### Primary: single-administration generalizability

For a one-time measure the relevant facet is **trials within one occasion**. The
coefficient is the generalizability of a person's score over that facet:

$$G = \frac{\text{Var}(person)}{\text{Var}(person) + \frac{\text{Var}(trial)}{i}}$$

estimated from `value ~ 1 + (1 | person)` on trial-level data restricted to the
baseline occasion. This is the same algebraic form as Equation 2 of Sliwinski et
al. (2018) with the *person* facet in the numerator instead of the occasion
facet — appropriate because a one-time measure is used to separate people, not
to track change.

Three companions are reported alongside it:

- **Split-half with Spearman-Brown**, the familiar form for reviewers.
- **SEM ÷ between-person SD.** The standard error of a single person's score
  relative to the spread across people. This is the most directly interpretable
  number for one-time use: at 0.4 an individual score is usable, at 0.9 it
  carries almost no individual-level information.
- **Trials required** to reach G = 0.70 and 0.80 at a single administration,
  by Spearman-Brown on the trial facet.

### Secondary: the EMA coefficients, and why they are not the answer here

Reported for completeness, following Sliwinski, Mogle, Hyun, Munoz, Smyth, &
Lipton (2018), *Assessment*, 25(1), 14–30
([doi:10.1177/1073191116643164](https://doi.org/10.1177/1073191116643164)):

$$\text{BP Reliability} = \frac{\text{Var}(BP)}{\text{Var}(BP) + \frac{\text{Var}(WP)}{n}}
\qquad
\text{WP Reliability} = \frac{\text{Var}(WP_{occasion})}{\text{Var}(WP_{occasion}) + \frac{\text{Var}(WP_{trial})}{i}}$$

`n` is the number of **assessments averaged per person** — not the sample size.
It reaches 56–70 in the original protocol and caps at 3 here.

Two cautions apply to this study:

- **WP Reliability is biased upward by the crossover.** `Var(WP_occasion)` is
  meant to be natural fluctuation; here it absorbs the treatment effect, so a
  working intervention inflates what looks like measurement reliability.
- **BP Reliability at `n` > 1 describes an averaged score**, which is not how
  these measures are being used.

### Task mapping

| Sliwinski et al. (2018) | This dataset | Trials/occasion |
|---|---|---|
| Symbol search — median RT of correct trials (12 trials) | Symbol Search — median RT of accurate trials | 18 |
| Dot memory — Euclidean-distance error score (2 trials) | Dot Memory — Euclidean-distance error score | 3 trials / 9 dots |
| N-back — proportion correct, odd/even split, `i` = 2 | Color Shapes — CorRec (HIT − FA), odd/even split, `i` = 2 | 10 → `i` = 2 |

Color Shapes takes the n-back handling because CorRec needs both change and
no-change trials, so a single trial cannot yield a score.

## Setup

```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(readr)
library(stringr)
library(tibble)
library(knitr)
library(lme4)
```

```{r}
raw_dir <- params$raw_dir
output_dir <- params$output_dir
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

timestamp <- format(Sys.time(), "%Y-%m-%d-T%H-%M-%S")

rm2c2_version <- numeric_version(read.dcf("DESCRIPTION", fields = "Version")[[1]])
packageVersion <- function(pkg) {
  if (identical(pkg, "RM2C2")) return(rm2c2_version)
  utils::packageVersion(pkg)
}

source(file.path("R", "add_to.R"))
source(file.path("R", "mult_by.R"))
source(file.path("R", "distance.R"))
source(file.path("R", "SDT_adj.R"))
source(file.path("R", "score_symbol_search.R"))
source(file.path("R", "score_dot_memory.R"))
source(file.path("R", "score_color_shapes.R"))
source(file.path("R", "sample_data_prep.R"))

prepared <- prepare_sample_data(raw_dir)
```

## Build Scores

```{r}
add_session <- function(d) mutate(d, session = as.numeric(str_split_i(occasion, "_", 2)))

symbol_scored <- prepared$symbol_raw |> score_symbol_search()
dot_scored    <- prepared$dot_raw |> score_dot_memory() |> ungroup()
color_scored  <- prepared$color_raw |> score_color_shapes()

# Unit-level tables (one row per trial, or per odd/even half for Color Shapes)
ss_unit <- symbol_scored |>
  filter(accuracy == 1, !is.na(response_time)) |>
  transmute(person = user_id, occasion = visit_id, trial = as.character(trial_num),
            value = response_time)

dm_unit <- dot_scored |>
  transmute(person = user_id, occasion = visit_id, trial = as.character(trial_num),
            value = (r1_distance + r2_distance + r3_distance) / 3)

# Dot-level alternative: the trial facet is 9 dots, not 3 trials.
dm_unit_dot <- dot_scored |>
  select(person = user_id, occasion = visit_id, trial_num,
         r1_distance, r2_distance, r3_distance) |>
  pivot_longer(starts_with("r"), names_to = "dot", values_to = "value") |>
  transmute(person, occasion, trial = paste0(trial_num, "_", dot), value)

cs_unit <- color_scored |>
  rename(person = user_id, occasion = visit_id) |>
  mutate(half = if_else(trial_num %% 2 == 1, "odd", "even")) |>
  group_by(person, occasion, half) |>
  summarise(value = sum(HIT) / sum(trial_type == 1) - sum(FA) / sum(trial_type == 0),
            .groups = "drop") |>
  rename(trial = half)
```

```{r}
ss_occ <- ss_unit |> group_by(person, occasion) |>
  summarise(score = median(value), .groups = "drop")
dm_occ <- dm_unit |> group_by(person, occasion) |>
  summarise(score = mean(value), .groups = "drop")
cs_occ <- color_scored |>
  rename(person = user_id, occasion = visit_id) |>
  group_by(person, occasion) |>
  summarise(score = sum(HIT) / sum(trial_type == 1) - sum(FA) / sum(trial_type == 0),
            .groups = "drop")

# Baseline = first occasion per participant (keeps 351 and 353).
baseline_ids <- add_session(ss_occ) |>
  group_by(person) |> slice_min(session, n = 1) |> ungroup() |>
  select(person, occasion)

kable(
  tibble(
    quantity = c("participants", "baseline occasions", "all occasions",
                 "participants with session == 1"),
    n = c(n_distinct(ss_occ$person), nrow(baseline_ids), nrow(ss_occ),
          n_distinct(add_session(ss_occ) |> filter(session == 1) |> pull(person)))
  ),
  caption = "Sample. Baseline takes the first occasion per participant."
)
```

## Primary: Single-Administration Reliability at Baseline

```{r}
# Method-of-moments person/trial decomposition - closed form, so the bootstrap
# below stays cheap.
g_components <- function(u) {
  per <- u |> group_by(person) |> summarise(i = n(), m = mean(value), .groups = "drop")
  N <- sum(per$i); P <- nrow(per); gm <- weighted.mean(per$m, per$i)
  MS_B <- sum(per$i * (per$m - gm)^2) / (P - 1)
  MS_W <- (u |> left_join(per, by = "person") |>
             summarise(s = sum((value - m)^2)) |> pull(s)) / (N - P)
  i0 <- (N - sum(per$i^2) / N) / (P - 1)
  Vp <- (MS_B - MS_W) / i0
  list(Vp = Vp, Vt = MS_W, i = mean(per$i),
       G = Vp / (Vp + MS_W / mean(per$i)))
}

split_half_sb <- function(u) {
  h <- u |>
    mutate(h = if_else(as.integer(factor(trial)) %% 2 == 1, "a", "b")) |>
    group_by(person, h) |> summarise(m = mean(value), .groups = "drop") |>
    pivot_wider(names_from = h, values_from = m)
  r <- cor(h$a, h$b, use = "complete.obs")
  2 * r / (1 + r)
}
```

```{r}
# trials_per_unit: Color Shapes units are 5-trial halves, the others are trials.
baseline_units <- list(
  "Symbol Search (median RT, accurate trials)" = list(u = ss_unit,     occ = ss_occ, tpu = 1),
  "Dot Memory (trial-level error)"             = list(u = dm_unit,     occ = dm_occ, tpu = 1),
  "Dot Memory (dot-level error)"               = list(u = dm_unit_dot, occ = dm_occ, tpu = 1),
  "Color Shapes (CorRec, odd/even halves)"     = list(u = cs_unit,     occ = cs_occ, tpu = 5)
) |>
  map(function(x) {
    x$u <- semi_join(x$u, baseline_ids, by = c("person", "occasion"))
    x$occ <- semi_join(x$occ, baseline_ids, by = c("person", "occasion"))
    x
  })

baseline <- imap_dfr(baseline_units, function(x, nm) {
  g <- g_components(x$u)
  need <- function(target) ceiling((target * g$Vt) / ((1 - target) * g$Vp)) * x$tpu
  tibble(
    task = nm,
    i = g$i,
    trials = g$i * x$tpu,
    Var_person = g$Vp,
    Var_trial = g$Vt,
    G = g$G,
    split_half_SB = split_half_sb(x$u),
    SD_between = sd(x$occ$score),
    SEM = sqrt(g$Vt / g$i),
    SEM_over_SD = sqrt(g$Vt / g$i) / sqrt(max(g$Vp, 0) + g$Vt / g$i),
    trials_for_0.70 = need(0.70),
    trials_for_0.80 = need(0.80)
  )
})

kable(baseline |> select(task, trials, G, split_half_SB, SEM, SD_between, SEM_over_SD),
      digits = 3,
      caption = "Single administration at baseline (n = 52). SEM_over_SD is the precision of one person's score relative to the between-person spread.")
```

```{r}
kable(baseline |> select(task, trials, `trials_for_0.70`, `trials_for_0.80`),
      caption = "Trials a single administration would need to reach these coefficients.")
```

### Bootstrap

```{r}
set.seed(20260906)
baseline_cache <- file.path(output_dir,
                            str_c("baseline_G_boot_", params$n_boot_baseline, ".rds"))

persons <- sort(unique(baseline_ids$person))
bsplit <- map(baseline_units,
              ~ split(.x$u |> select(person, value), .x$u$person)[as.character(persons)])

baseline_boot <- if (file.exists(baseline_cache)) readRDS(baseline_cache) else
  map_dfr(seq_len(params$n_boot_baseline), function(b) {
    j <- sample(seq_along(persons), length(persons), replace = TRUE)
    tag <- seq_along(j)
    imap_dfr(bsplit, function(s, nm) {
      u <- bind_rows(map2(s[j], tag, ~ mutate(.x, person = .y)))
      tibble(task = nm, G = g_components(u)$G)
    })
  })

if (!file.exists(baseline_cache)) saveRDS(baseline_boot, baseline_cache)

baseline_ci <- baseline_boot |>
  group_by(task) |>
  summarise(`G 95% CI` = sprintf("[%.2f, %.2f]", quantile(G, .025), quantile(G, .975)),
            .groups = "drop")

kable(baseline |> select(task, G) |> left_join(baseline_ci, by = "task"), digits = 3,
      caption = paste0("Percentile bootstrap over participants, B = ", params$n_boot_baseline, "."))
```

### Alternative dependent variables for the weak tasks

```{r}
alt_dvs <- list(
  "Color Shapes - CorRec (HIT-FA), halves"   = list(u = cs_unit, tpu = 5),
  "Color Shapes - trial-level accuracy"      = list(
    u = color_scored |> transmute(person = user_id, occasion = visit_id,
                                  trial = as.character(trial_num),
                                  value = HIT + CR), tpu = 1),
  "Color Shapes - proportion correct, halves" = list(
    u = color_scored |> rename(person = user_id, occasion = visit_id) |>
      mutate(h = if_else(trial_num %% 2 == 1, "odd", "even")) |>
      group_by(person, occasion, h) |> summarise(value = mean(HIT + CR), .groups = "drop") |>
      rename(trial = h), tpu = 5)
) |> map(function(x) { x$u <- semi_join(x$u, baseline_ids, by = c("person", "occasion")); x })

kable(imap_dfr(alt_dvs, function(x, nm) {
  g <- g_components(x$u)
  tibble(DV = nm, i = g$i, G = g$G,
         trials_for_0.80 = ceiling((0.80 * g$Vt) / ((1 - 0.80) * g$Vp)) * x$tpu)
}), digits = 3, caption = "Color Shapes under alternative scoring, single administration.")
```

## Secondary: EMA Coefficients Across All Occasions

The two tables below apply Equations 1 and 2 to all 155 occasions. **They do not
answer this study's question.** Equation 2 in particular is inflated by the
crossover: with an intervention between sessions, `Var(WP_occasion)` contains the
treatment effect rather than natural fluctuation.

```{r}
vc_person_occasion <- function(d) {
  # d: one row per OCCASION (occasion-level score). Residual is Var(WP_occasion).
  m <- lmer(score ~ 1 + (1 | person), data = d, REML = TRUE)
  v <- as.data.frame(VarCorr(m))
  list(BP = v$vcov[v$grp == "person"], WP = v$vcov[v$grp == "Residual"])
}
vc_occasion_trial <- function(d) {
  # d: one row per TRIAL. Residual is the trial facet; person:occasion is the
  # within-person occasion facet.
  m <- lmer(value ~ 1 + (1 | person) + (1 | person:occasion), data = d, REML = TRUE)
  v <- as.data.frame(VarCorr(m))
  list(WPocc = v$vcov[v$grp == "person:occasion"], WPtrial = v$vcov[v$grp == "Residual"])
}
bp_reliability <- function(BP, WP, n) BP / (BP + WP / n)
wp_reliability <- function(WPocc, WPtrial, i) WPocc / (WPocc + WPtrial / i)

ema_tasks <- list(
  "Symbol Search (median RT, accurate trials)" = list(occ = ss_occ, trial = ss_unit),
  "Dot Memory (Euclidean error distance)"      = list(occ = dm_occ, trial = dm_unit),
  "Color Shapes (CorRec = HIT - FA)"           = list(occ = cs_occ, trial = cs_unit)
)

ema <- imap_dfr(ema_tasks, function(t, nm) {
  a <- vc_person_occasion(t$occ); b <- vc_occasion_trial(t$trial)
  i <- mean(t$trial |> count(person, occasion) |> pull(n))
  tibble(task = nm, i = i,
         `ICC (n = 1)` = bp_reliability(a$BP, a$WP, 1),
         `BP rel (n = 3)` = bp_reliability(a$BP, a$WP, 3),
         `WP rel (contaminated)` = wp_reliability(b$WPocc, b$WPtrial, i))
})

kable(ema, digits = 3,
      caption = "Equations 1 and 2 across all occasions. WP is confounded with the intervention.")
```

```{r}
retest <- imap_dfr(ema_tasks, function(t, nm) {
  w <- add_session(t$occ) |> filter(session %in% 1:2) |>
    select(person, session, score) |>
    pivot_wider(names_from = session, values_from = score, names_prefix = "s")
  tibble(task = nm, n_pairs = sum(complete.cases(w)),
         `r(session 1, 2)` = cor(w$s1, w$s2, use = "complete.obs"))
})

kable(retest, digits = 3,
      caption = "Session 1 to 2 correlation. NOT test-retest: session 2 is post-intervention.")
```

## Condition-Adjusted Analysis

Supplying a randomization file at `params$condition_file` (a CSV with `person`,
`session`, and `condition`) lets the treatment effect be modelled as a fixed
effect, so the residual occasion variance is interpretable as fluctuation rather
than as the intervention working.

```{r}
if (nzchar(params$condition_file) && file.exists(params$condition_file)) {
  conditions <- read_csv(params$condition_file, show_col_types = FALSE) |>
    mutate(person = as.character(person), session = as.numeric(session))

  cond_adj <- imap_dfr(ema_tasks, function(t, nm) {
    u <- add_session(t$trial) |>
      left_join(conditions, by = c("person", "session")) |>
      filter(!is.na(condition))
    m <- lmer(value ~ condition + (1 | person) + (1 | person:occasion), data = u, REML = TRUE)
    v <- as.data.frame(VarCorr(m))
    i <- mean(u |> count(person, occasion) |> pull(n))
    tibble(task = nm, i = i,
           `WP rel (condition-adjusted)` = wp_reliability(
             v$vcov[v$grp == "person:occasion"], v$vcov[v$grp == "Residual"], i))
  })
  kable(cond_adj, digits = 3,
        caption = "Within-person reliability net of the treatment effect.")
} else {
  cat("No condition file supplied. Set params$condition_file to the randomization CSV",
      "(columns: person, session, condition) to run this section.")
}
```

## Write Output Files

```{r}
baseline_path <- file.path(output_dir, str_c("onetime_reliability_", timestamp, ".csv"))
ema_path <- file.path(output_dir, str_c("ema_reliability_", timestamp, ".csv"))

baseline_out <- baseline |> left_join(baseline_ci, by = "task")
write_csv(baseline_out, baseline_path)
write_csv(ema, ema_path)

kable(tibble(
  file_type = c("one_time_reliability", "ema_reliability"),
  file_path = c(baseline_path, ema_path)
))
```

## Conclusions

```{r}
kable(baseline_out |> select(task, trials, G, `G 95% CI`, SEM_over_SD, trials_for_0.80),
      digits = 3)
```

**Symbol Search is adequate as a one-time measure.** G ≈ 0.86 with an SEM about
0.4 of the between-person SD. Eighteen RT trials is already more than the twelve
needed for 0.80, so it is doing what a one-time measure needs to do.

**Dot Memory is marginal, and should be scored at the dot level.** Treating the
trial facet as 9 dots rather than 3 trials raises G from ≈0.42 to ≈0.60 on
exactly the same data — the current trial-level score discards usable
information. Even so it falls short of 0.80, which would need roughly 8 trials
rather than 3. Usable for group comparisons, not for individual-level inference.

**Color Shapes should not be used as a one-time individual-difference measure.**
G ≈ 0.14, and the SEM is ~96% of the between-person SD: a single score is nearly
as uncertain as the entire spread across participants. No dependent variable
rescues it — proportion correct is the best of the three tried and still only
reaches ≈0.25. The cause is structural: between-person variance is small
relative to binomial noise at 10 trials, so reaching 0.80 would take on the order
of 120+ trials. Reporting a low coefficient and proceeding would be worse than
stating plainly that one administration cannot support person-level conclusions.

**The general point.** These tasks are short by design because the EMA protocol
recovers reliability by averaging ~56 occasions — Equation 1 does that work, and
at `n = 56` even a weak single-occasion score reaches 0.97. Used once, each task
is left with only what one short block supports. Symbol Search has enough trials
to survive the change of context; the other two do not.

### Caveats

- Bootstrap intervals reflect participant sampling only, not trial-level
  resampling within a participant.
- No transformations were applied, following the original study. RT and error
  distance are both right-skewed; a log transform is worth testing before
  publication.
- Odd/even splitting a 10-trial task does not preserve Color Shapes' 5/5
  change balance — halves carry 1–4 change trials, so CorRec takes only 13
  distinct values at half length and 79 of 310 halves sit at ceiling.
- The baseline analysis assumes session 1 is genuinely pre-intervention for all
  participants, including 351 and 353 whose first occasion is labelled session 2.
